Probability

### 2016-04-18 prob.tex
restructure quantum prob

\documentclass{article}
\usepackage{fullpage}
\usepackage{bbm}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{multicol}

\theoremstyle{remark}
\newtheorem{example}{Example}
\newcommand{\events}{\ensuremath{\mathcal{F}}}
\newcommand{\qevents}{\ensuremath{\mathcal{A}}}
\newcommand{\pmeas}{\ensuremath{\mu}}
\newcommand{\Hilb}{\mathcal{H}}
\newcommand{\pr}[2]{\langle #1, #2 \rangle}
\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\bra}[1]{\langle#1|}
\newcommand{\ip}[2]{\langle #1 | #2 \rangle}
\newcommand{\proj}[1]{|#1 \rangle\langle #1 |}

@@

A classical model decides the occurrence or non-occurence of all
events simultaneously which is inconsistent with quantum
mechanics. Indeed, in the quantum world, there are (non-commuting)
events which cannot happen simultaneously. To accommodate this
situation, we completely abandon the sample space $\Omega$ and define
and reason directly about events. Thus a quantum probability space
will consist of just two components: a set of events $\qevents$ and a
probability measure $\phi : \qevents \rightarrow [0,1]$. These
components are defined as follows.

We first assume an ambient Hilbert space $\Hilb$ and define the set of
events $\qevents$ as \emph{projections} on $\Hilb$. Similarly to the
classical case, a probability measure is a function
$\phi : \qevents \rightarrow [0,1]$ satisfying:
\begin{itemize}
\item $\phi(\mathbbm{1}) = 1$, and
\item for all $A \in \qevents$, we have $\phi(A^*A) \geq 0$.
\end{itemize}

As an example, let $P_1, P_2, \ldots, P_k$ be mutually orthogonal
projections on $\Hilb$ with sum $\mathbbm{1}$ and define the event
space $\qevents$ to be the linear span of these operators:
\[
\qevents = \{ \sum_{j=1}^k \lambda_jP_j ~|~ \lambda_i, \ldots, \lambda_k \in \mathbb{C} \}. 
\]
Each state $\ket{\psi}$ of the Hilbert space induces a probability
measure $\phi_\psi : \qevents \rightarrow [0,1]$ defined as follows:
\[
\phi_\psi(A) = \ip{\psi}{A\psi}
\]

%% If we want to follow Swart~\cite{Swart2013} 
%% I am following sec 4 of https://www.math.ru.nl/~maassen/lectures/Trieste.pdf

Concrete example: consider the two qubit Hilbert space with
computational bases $\ket{0}$ and $\ket{1}$ and consider
the following families of projections:
\begin{itemize}
\item Family I: $\proj{0}$, $\proj{1}$
\item Family II: $\proj{+}$, $\proj{-}$
\end{itemize}
and consider the two states \ldots

>>>>

\documentclass{article}
\usepackage{fullpage}
\usepackage{bbm}
\usepackage{bbold}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{multicol}

\theoremstyle{remark}
\newtheorem{example}{Example}
\newcommand{\events}{\ensuremath{\mathcal{A}}}
\newcommand{\qevents}{\ensuremath{\mathcal{A}}}
\newcommand{\pmeas}{\ensuremath{\mu}}
\newcommand{\Hilb}{\mathcal{H}}
\newcommand{\pr}[2]{\langle #1, #2 \rangle}
\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\bra}[1]{\langle#1|}
\newcommand{\ip}[2]{\langle #1 | #2 \rangle}
\newcommand{\proj}[1]{|#1 \rangle\langle #1 |}
\newcommand{\ps}{\texttt{+}}
\newcommand{\ms}{\texttt{-}}

@@

The mathematical framework above assumes that one has complete
knowledge of the events and their relationships. But even in many
classical situations, the structure of the event space is only
partially known and the precise dependence of two events on each other
cannot be determined with certainty. In the quantum case, this partial
knowledge is compounded by the fact that not all quantum events can be
observed simultaneously.  Indeed, in the quantum world, there are
non-commuting events which cannot even happen simultaneously. To
accommodate these more complex situations, we completely abandon the
sample space~$\Omega$ and define and reason directly about
events. Thus a quantum probability space will consist of just two
components: a set of events $\qevents$ and a probability measure
$\mu : \qevents \rightarrow [0,1]$. We give an example before giving
the formal definition.

Consider the two-qubit Hilbert space with computational basis
$\ket{0}$ and $\ket{1}$ and states:
\[
\ket{\ps}=\frac{\ket{0}+\ket{1}}{\sqrt{2}}, \qquad\ket{\ms}=\frac{\ket{0}-\ket{1}}{\sqrt{2}}\,.
\]
A possible space of events is the set
$\{ \mathbb{0}, \proj{0}, \proj{1}, \mathbb{1}\}$ consisting of the
projection operators $\proj{0}$ and $\proj{1}$ as well as the empty
projection $\mathbb{0}$ and the unit projection
$\mathbb{1} = \proj{0}+\proj{1}$. Each event is interpreted as a
possible post-measurement of a quantum system as follows: given some
arbitrary current quantum $\ket{\psi}$ to be measured, the event
$\proj{0}$ states that the post-measurement state will be $\ket{0}$;
the event $\proj{1}$ states that the post-measurement state will be
$\ket{1}$; the event $\mathbb{1}$ states that the post-measurement
state will be a linear combination of $\ket{0}$ and $\ket{1}$; and the
event $\mathbb{0}$ states that the post-measurement state will be the
empty state. Clearly irrespective of the current state $\ket{\psi}$
and irrespective of the particular experiment, the probability of
event $\mathbb{0}$ will always be 0 (it is an impossible event) and
the probability of event $\mathbb{1}$ will always be 1 (it is a
certain event). The probabilities attached to other events will depend
on the particular state in question. If the state is $\ket{0}$, the
probability of event $\proj{0}$ is 1 and the probability of event
$\proj{1}$ is 0. If the state is $\ket{\ps}$, the probability of both
events $\proj{0}$ and $\proj{1}$ will be $\frac{1}{2}$. The same
Hilbert space setup admits other possible spaces of events. For
example, another space of events could be
$\{ \mathbb{0}, \proj{\ps}, \proj{\ms}, \mathbb{1}\}$. Using that
space of events, the state $\ket{0}$ maps both events $\proj{\ps}$ and
$\proj{\ms}$ to the probability $\frac{1}{2}$.

% \begin{itemize}
% \item Family I: $\proj{0}$, $\proj{1}$ 
% \item Family II: $\proj{\ps}$, $\proj{\ms}$
% \end{itemize}
% In family I, all operators can be expressed as $\left\{ \lambda_{1}\proj{0}+\lambda_{2}\proj{1}~\middle|~\lambda_{1},\lambda_{2}\in\mathbb{C}\right\} $.
% In order to identify projectors among them, we need to solve the following
% two equations.
% \begin{eqnarray*}
% \lambda_{1}\proj{0}+\lambda_{2}\proj{1} & = & \left(\lambda_{1}\proj{0}+\lambda_{2}\proj{1}\right)^{*}=\lambda_{1}^{*}\proj{0}+\lambda_{2}^{*}\proj{1}\\
% \lambda_{1}\proj{0}+\lambda_{2}\proj{1} & = & \left(\lambda_{1}\proj{0}+\lambda_{2}\proj{1}\right)^{2}=\lambda_{1}^{2}\proj{0}+\lambda_{2}^{2}\proj{1}
% \end{eqnarray*}
% Therefore, we actually have $\lambda_{1},\lambda_{2}\in\left\{ 0,1\right\} $,
% and there are only four projections: $0$, $\proj{0}$, $\proj{1}$,
% and $\mathbb{1}=\proj{0}+\proj{1}$. Then, $\mu_{\ket{\ps}}$ and
% $\mu_{\ket{\ms}}$ of each projections are 
% \[
% \begin{array}{c@{\qquad\qquad}c}
% \begin{array}{rcl}
% \mu_{\ket{\ps}}(0) & = & 0\\
% \mu_{\ket{\ps}}(\proj{0}) & = & 1/2\\
% \mu_{\ket{\ps}}(\proj{1}) & = & 1/2\\
% \mu_{\ket{\ps}}(\mathbb{1}) & = & 1
% \end{array} & \begin{array}{rcl}
% \mu_{\ket{\ms}}(0) & = & 0\\
% \mu_{\ket{\ms}}(\proj{0}) & = & 1/2\\
% \mu_{\ket{\ms}}(\proj{1}) & = & 1/2\\
% \mu_{\ket{\ms}}(\mathbb{1}) & = & 1
% \end{array}\end{array}
% \]
% Similarly, $\mu_{\ket{\ps}}$ and $\mu_{\ket{\ms}}$ gives two probability
% for family II as well.
% \[
% \begin{array}{c@{\qquad\qquad}c}
% \begin{array}{rcl}
% \mu_{\ket{\ps}}(0) & = & 0\\
% \mu_{\ket{\ps}}(\proj{+}) & = & 1\\
% \mu_{\ket{\ps}}(\proj{\ms}) & = & 0\\
% \mu_{\ket{\ps}}(\mathbb{1}) & = & 1
% \end{array} & \begin{array}{rcl}
% \mu_{\ket{\ms}}(0) & = & 0\\
% \mu_{\ket{\ms}}(\proj{\ps}) & = & 0\\
% \mu_{\ket{\ms}}(\proj{\ms}) & = & 1\\
% \mu_{\ket{\ms}}(\mathbb{1}) & = & 1
% \end{array}\end{array}
% \]

We now formalize a \emph{quantum probability space} as
follows~\cite{Maassen2010,Swart2013}. We first assume an ambient
Hilbert space $\Hilb$ and define the set of events $\qevents$ as all
linear combinations of \emph{projections} on $\Hilb$. Specifically,
let $P_1, P_2, \ldots, P_k$ be mutually orthogonal projections on
$\Hilb$ with sum $\mathbb{1}$ and define the event space $\qevents$ to
be the linear span of these operators:
\[
\qevents=\left\{ \sum_{j=1}^{k}\lambda_{j}P_{j}~\middle|~\lambda_{1},\ldots,\lambda_{k}\in\mathbb{C}\right\} .
\]
Each quantum state $\ket{\psi}$ induces a probability measure
$\mu_\psi : \qevents \rightarrow [0,1]$ on the space of events
defined as follows:
\[
\mu_\psi(A) = \ip{\psi}{A\psi}
\]
Similarlly to the classical case, this probability measure satisfies:
\begin{itemize}
\item $\mu(\mathbb{1}) = 1$, and
\item for all $A \in \qevents$, we have $\mu(A^*A) \geq 0$.
\end{itemize}

% \yutsung{If we follow \cite{Maassen2010,Swart2013}, then we also
% need 
% \begin{itemize}
% \item $\mu$ can be extended to a linear functional~$\mu:alg\left(\qevents\right)\rightarrow\mathbb{C}$,
% where $alg\left(\qevents\right)$ is the minimal $*$-algebra generated
% by $\qevents$.
% \end{itemize}
% Also, 
% \begin{itemize}
% \item for all $A\in\qevents$, we have $\mu(A^{*}A)\geq0$. 
% \end{itemize}
% should be 
% \begin{itemize}
% \item for all $A\in alg\left(\qevents\right)$, we have $\mu(A^{*}A)\geq0$. 
% \end{itemize}
% because we have $\mu(A^{*}A)=\mu(A^{2})=\mu(A)$ if $A$ is a projection.}\\

% \yutsung{$\left\{ \sum_{j=1}^{k}\lambda_{j}P_{j}~\middle|~\lambda_{1},\ldots,\lambda_{k}\in\mathbb{C}\right\} $
% is the minimal $*$-algebra generated by $P_{1},P_{2},\ldots,P_{k}$,
% but it contains all possible observables $P_{1},P_{2},\ldots,P_{k}$
% can generate (and something more) not just projections. For example,
% $2\mathbb{1}\in\left\{ \sum_{j=1}^{k}\lambda_{j}P_{j}~\middle|~\lambda_{1},\ldots,\lambda_{k}\in\mathbb{C}\right\} $.}\\

% \yutsung{So $\mu_{\psi}$ maps the projections generated by $P_{1},P_{2},\ldots,P_{k}$
% to $[0,1]$, and maps $\left\{ \sum_{j=1}^{k}\lambda_{j}P_{j}~\middle|~\lambda_{1},\ldots,\lambda_{k}\in\mathbb{C}\right\} $
% to $\mathbb{C}$...}\\
